Synthesis report

### 2020-04-27 analysis/reports/fia_size.Rmd
LOOK AT THIS NONSENSE

The small datasets are basically all FIA.

Note that the color scale is log transformed, so the largest communities have e^`r max(all_di$log_nparts)`, or `r exp(max(all_di$log_nparts))`, elements in the feasible set!

# 95 interval vs nparts

```{r 95 interval}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio_2t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio_2t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

@@

```{r 95 ratio v nparts scatter}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log_nparts, y = skew_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = simpson_95_ratio_2t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

@@

```{r percentiles v 95 interval}

ggplot(filter(all_di, !singletons), aes(x = skew_95_ratio_2t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_95_ratio_2t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

# Binned by ranked nparts and ranked 95 intervals

```{r binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_2t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_2t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

# 95 interval vs nparts

```{r 1t 95 interval}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

```{r 1t 95 ratio v nparts scatter}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log_nparts, y = skew_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = simpson_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

```{r 1t percentiles v 95 interval}

ggplot(filter(all_di, !singletons), aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

# Binned by ranked nparts and ranked 95 intervals

```{r 1t binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r 1t binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r 1t relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

```{r overall relationships}

ggplot(filter(all_di, !singletons), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()

```{r overall relationships by da t}

ggplot(filter(all_di, !singletons, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()+
  facet_wrap(vars(dat), scales = "free_y")

```{r gentry}

ggplot(filter(all_di, dat == "gentry", !singletons), aes(x = log(n0/s0), y = simpson_percentile)) +
  geom_point()

gentry_cat <- filter(all_di, dat == "gentry", !singletons) %>%
  mutate(low_avgn = log(n0/s0) < 1.25)

ggplot(gentry_cat, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y")

>>>>

## Resrict to sizes found in FIA

```{r fia limits}

fia_max_s <- max(filter(all_di, !singletons, dat == "fia_short")$s0)
fia_max_n <- max(filter(all_di,!singletons,  dat == "fia_short")$n0)

fia_sized <- filter(all_di, s0 <= fia_max_s, !singletons, n0 <= fia_max_n) %>%
  mutate(fia_yn = dat == "fia_short")

ggplot(fia_sized, aes(x = log(s0), y= log(n0), color = dat)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d()

mean(fia_sized$dat == "fia_short")

@@

95% of the datasets with state variables comparable to the FIA ranges are FIA.

```{r fia sized hists}

ggplot(filter(fia_sized, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram()

ggplot(filter(fia_sized, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")

ggplot(filter(fia_sized, s0 > 2), aes(x = simpson_percentile)) +
  geom_histogram()

ggplot(filter(fia_sized, s0 > 2), aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(fia_yn), scales = "free_y")

@@

```{r on map}

ggplot(fia_sized, aes(x = log(s0), y= log(n0), color = skew_percentile)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c()

ggplot(fia_sized, aes(x = log(s0), y= log(n0), color = simpson_percentile)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c()
```
